### 2016-03-08 article.Rnw
edits based on comments of rbs

\documentclass[article, shortnames]{jss}
\usepackage[utf8]{inputenc} %unicode support
\usepackage{booktabs}

\author{Eduard Sz\"ocs\\Universit\"at Koblenz-Landau \And 
        Ralf B. Sch\"afer\\Universit\"at Koblenz-Landau}
\title{\pkg{webchem}: An \proglang{R} Package to retrieve Chemical Information from the Web}

\Plainauthor{Eduard Sz\"ocs, Ralf B. Sch\"afer}
\Plaintitle{webchem: an R Package to retrieve Chemical Information from the Web}
\Shorttitle{\pkg{webchem}: Chemical Information from the Web}

\Abstract{
A lot of chemical information is freely available online, including identifiers, experimental and predicted chemical properties.
However, these data are scattered over various data sources and not easily accessible to researchers.
Manual searching and downloading of such data is time-consuming and error-prone.

We developed the open-source \proglang{R} package \pkg{webchem} that allows users to automatically query chemical data from currently 11 web-sources. 
These cover a broad spectrum of information.
The data are automatically imported into an \proglang{R} object and can directly be used in subsequent analyses.

\pkg{webchem} enables easy, structured and reproducible data retrieval and usage from publicly available web-sources.
In addition, it facilitates data cleaning, identification and reporting of substances.
Consequently, it reduces the time researchers need to spend on chemical data compilation.
}
\Keywords{ecotoxicology, chemistry, data-cleaning, web-scraping, ropensci}
\Plainkeywords{ecotoxicology, chemistry, data-cleaning, web-scraping, ropensci}

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{2012-06-04}
%% \Acceptdate{2012-06-04}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Eduard Sz\"ocs\\
  Institute for Environmental Sciences\\
  Universit\"at Koblenz-Landau\\
  Fortstraße 7\\
  76829 Landau, Germany\\
  E-mail: \email{szoecs@uni-landau.de}\\
  URL: \url{https://edild.github.io}
}

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

\begin{document}

<<global_options, include=FALSE>>=
library(knitr)
knitr::opts_chunk$set(eval = TRUE, cache = TRUE)
knitr::render_sweave()
knitr::opts_chunk$set(prompt = TRUE,
                      comment = NA,
                      highlight = FALSE,
                      dev.args = list(pointsize = 11),
                      # fig.width = 4.9,
                      # fig.height = 3.675,
                      fig.align = 'center',
                      out.format = 'sweave',
                      tidy = FALSE,
                      fig.path = '')
# 
hook.c <- function(x, options) {
  paste0('\\begin{CodeChunk}\n', x, '\\end{CodeChunk}')
}
hook.i <- function(x, options) {
  paste0(c('\\begin{CodeInput}', knitr:::hilight_source(x, 'latex', options), '\\end{CodeInput}', ''), collapse = '\n')
}
hook.s <- function(x, options) {
  paste0('\\begin{CodeOutput}\n', x, '\\end{CodeOutput}\n')
}
knitr::knit_hooks$set(chunk = hook.c,
               source = hook.i,
               output = hook.s,
               message = hook.s,
               warning = hook.s,
               plot = knitr::hook_plot_tex)

options(prompt = "R> ", continue = "+  ", width = 80, useFancyQuotes = FALSE)
@

\section[Introduction]{Introduction}
Before each statistical analysis, data cleaning is often required to ensure a good data quality.
Data cleaning is the process of detecting errors and inconsistencies in data sets \citep{Chapman_2005}.
In practice, the data cleaning step is often more time consuming than the subsequent statistical analysis, particularly, when the analysis relies on multiple data sources that need to be joined.

When dealing with chemical datasets (e.g. environmental monitoring data, toxicological data), a first step is often to validate the names of chemicals or to link them to unique codes that simplify subsequent querying and appending of compound-related physico-chemical or toxicological information.
Several web sources provide chemical names or link them to unique codes (see also section \emph{Data sources} below).
However, manual searching for each compound, often through a graphical web interface, is tedious, error-prone and not reproducible \citep{Peng_2009}.

To simplify, robustify and automate this task, i.e. to search and retrieve chemical information from the web, we created the \pkg{webchem} package for the free and open source \proglang{R} language \citep{r_2015, Wehrens_2011}.
\proglang{R} is one of the most widely used software for data cleaning, analysing and visualising data, and supports full reproducibility of each step \citep{Marwick_2016}.

We describe the basic functionality of the package and demonstrate with a few use cases how to clean and retrieve new data with \pkg{webchem}.

\section[Implementation and design details]{Implementation and design details}
The \pkg{webchem} package is written entirely in \proglang{R} and available under a MIT license.
The development repository is hosted on GitHub (\url{https://github.com/ropensci/webchem}) and a stable version is released on the official \proglang{R} repository (CRAN, \url{https://goo.gl/OKbZaF}).
\pkg{webchem} is part of the rOpenSci project (\url{https://ropensci.org/}), which aims at fully reproducible data analysis.

\pkg{webchem} follows best practices for scientific software \citep{wilson_best_2014, poisot_best_2015}, namely: (i) a public available repository with easy collaboration and an issue tracker (via GitHub), (ii) a non-restrictive license, version control (git), (iii) an elaborate test-suite covering more than 90\% of the relevant lines of code (currently \textasciitilde 1500 lines, using \pkg{testthat} \citep{wickham_testthat:_2011}), (iv) continuous integration (via Travis-CI and AppVeyor; testing on Linux \& Windows with current and development \proglang{R} versions), (v) in-source documentation (using \pkg{roxygen2} \citep{wickham_roxygen2:_2015}) and (vi) compliance with a style guide \citep{wickham_advanced_2015}.

\pkg{webchem} builds on top of the following \proglang{R} packages:
\pkg{RCurl} \citep{lang_rcurl:_2015} and \pkg{httr} \citep{wickham_httr} for data transfer,
\pkg{stringr} \cite{wickham_stringr:_2015} for string handling,
\pkg{xml2} \citep{wickham_xml2} and \pkg{rvest} \citep{wickham_rvest} for parsing HTML and XML,
\pkg{jsonlite} \citep{ooms_jsonlite_2014} for parsing JSON,
\pkg{rcdk} \citep{guha_rcdk} for parsing SMILES.
For parsing molfiles we use a lightweight implementation of \citep{Grabner_Varmuza_Dehmer_2012}.

Some data sources provide application programming interfaces (API).
Web APIs define functions that allow accessing services and data via http and return data in a specific way.
\pkg{webchem} uses the API of a data source provider, where this is provided.
For sources where an API is lacking, data is directly searched and extracted from the web pages, analogous to manual interaction with a website.

Only few design decisions have been made:
Each function name has a prefix and suffix separated by an underscore \citep{Chamberlain_Szocs_2013}.
They follow the format of \code{source\_function}, e.g. \code{cs\_compinfo} uses ChemSpider as source (see next section) to retrieve compound information.
Some functions require querying first a unique identifier from the data source and then use this identifier to query further information.
The prefix \code{get} is used to denote these functions, e.g. \code{get\_csid} to retrieve the identifier used in ChemSpider.

\pkg{webchem} is friendly to the resources of data providers. 
Between each request there is a time-out of 0.3 to 2 seconds depending on data-source. 
Therefore, processing of larger datasets can take some time, but still represents a major improvement compared to manual lookup.
We provide a link to the \emph{Terms of Use} of data-providers in the documentation of each function and we encourage the users to read these before using \pkg{webchem}.
Moreover, all functions return an URL of the source, which can be used for \mbox{(micro-)attribution}.

\section[Data sources]{Data sources}
The backbone of \pkg{webchem} are data sources providing their data and functionality to the public.
Currently, data can be retrieved from 11 sources.
These cover a broad spectrum of available data, like identifiers, experimental and predicted properties and legal information (Figure~\ref{fig:fig1}, a detailed overview of all sources is included as supplement):

\begin{description}
  \item[NIH Chemical Identifier Resolver (CIR) \citep{cir}]{A web service that converts from and to various chemical identifiers.}
  \item[Chemical Translation Service (CTS) \citep{Wohlgemuth_Haldiya_Willighagen_Kind_Fiehn_2010}]{A web service that converts from and to various chemical identifiers.}
  \item[ETOX \citep{etox}]{Information System Ecotoxicology and Environmental Quality Targets by the German Federal Environmental Agency. Provides basic identifiers, synonyms, ecotoxicological data and quality targets for different countries.}
  \item[PAN Pesticide Database \citep{pan}]{Information on pesticides - provides basic identifiers, ecotoxicological data and chemical properties.}
  \item[SRC Physprop \citep{physprop}]{Contains physical properties for over 41,000 chemicals.
  Physical properties collected from a wide variety of sources including experimental and modelled values.}
  \item[PubChem \citep{Kim_2016}]{PubChem is a public repository for information on chemical substances, providing identifiers, properties and synonyms.
  We use an interface to the PUG-REST web-service \citep{Kim_Thiessen_Bolton_Bryant_2015}.}
  \item[Wikidata \citep{wiki}]{Wikipedia contains information entries for over 15,000 chemicals \citep{Ertl_Patiny_Sander_Rufener_Zasso_2015}. Currently \pkg{webchem} can query only chemical identifiers.}
  \item[Compendium of Pesticide Common Names \citep{wood}]{The compendium provides information on pesticide common names, identifiers and classification.}
  \item[ChemID\emph{plus} \citep{Tomasulo_2002}]{is a large web-based database provided by the National Library of Medicine. It provides identifiers, synonyms, toxicological data and chemical properties.}
  \item[ChemSpider \citep{pence_chemspider:_2010}]{ChemSpider is a free chemical structure database providing access to over 40 million structures. It provides identifiers, properties and can also be used to convert identifiers.}
  \item[OPSIN  \citep{Lowe_Corbett_Murray-Rust_Glen_2011}]{The Open Parser for Systematic IUPAC nomenclature is a chemical name interpreter and provides InChI and SMILES identifiers.}
\end{description}

Though the data sources exhibit some overlap in the provided information, each has been selected because it provides unique information and we encourage the interested reader to consult the related source for details. 
However, we provide a brief overview in the Supporting Information.

\begin{figure}[ht]
  \centering
  \includegraphics{fig1.pdf}
  \caption{Overview of current data sources. Input and output possibilities currently implemented in the package.}
  \label{fig:fig1}
\end{figure}

\section[Use Cases]{Use Cases}
\subsection[Install webchem]{Installation}
\pkg{webchem} can be easily installed and loaded from CRAN:

<<install, eval=FALSE>>=
install.packages("webchem")
library(webchem)
@

<<load, echo=FALSE, eval=TRUE>>=
library(webchem)
library(ggplot2)
library(xtable)
@

The package is under active development and the latest development version is available from github and also permanently available at \url{http://dx.doi.org/10.5281/zenodo.46930}.
This document has been created using \pkg{webchem} version 0.1.

\subsection[Sample datasets]{Sample datasets}
To demonstrate the capabilities of  \pkg{webchem} we use two small publicly available real world datasets.
The datasets are only for demonstration purposes, have been slightly preprocessed (not shown) and are available through the package.

(i) \code{jagst}: This dataset comprises environmental monitoring data of organic substances in the river Jagst of the year 2013, Germany.
The data is publicly available and can be retrieved from \citep{lubw_2016}.
It comprises concentrations  (in $\mu g~/~L$) of  34 substances  on 13 sampling occasions.

<<loadjagst>>=
# load jagst dataset
data(jagst)
# print first 6 lines of data
head(jagst)
@

This dataset identifies substances only by substance names. Values below LOQ are indicated by a qualifier column.

(ii) \code{lc50}: This data consists of median acute lethal concentration for the water flea daphnia magna in 48 h tests ($LC_{50, D.magna, 48h}$) of 124 insecticides.
The data has been retrieved from the EPA ECOTOX database \citep{epa_2016}.

<<load_lc50>>=
data(lc50)
head(lc50)
@

This data set identifies the substances only by CAS numbers.

\subsection[Query Identifiers]{Query Identifiers}
The \code{jagst} data set covers 34 substances that are identified by (German) names.
Merging and linking these to other tables is hampered by differences and ambiguity in compound names.

One possibility to resolve this, is to use different chemical identifiers allowing easy identification.
There are several identifiers available, e.g.  registry numbers like CAS or EC, database identifiers like PubChemCID \citep{Kim_2016} or ChemSpiderID \citep{pence_chemspider:_2010}, line notations like SMILES \citep{Weininger_1990}, InChI and InChiKey \citep{Heller_McNaught_Pletnev_Stein_Tchekhovskoi_2015}. 
In this first example we query several identifiers to create a table that can be used as (i) supplemental information to a research article or (ii) to facilitate subsequent matching with other data.

As were are dealing with German substance names we start to query ETOX for CAS registry numbers.
A common work flow when dealing with web resources is to 1) query a unique identifier of the source, 2) use this identifier to retrieve additional information and 3) extract the parts that are needed from the \proglang{R}-object \citep{Chamberlain_Szocs_2013}.

First we search for ETOX internal ID numbers:

<<get_etoxid, message=FALSE>>=
# unique substance names
subs <- unique(jagst$substance)
# search ETOX IDs
ids <- get_etoxid(subs, match = 'best')
head(ids)
@

Only three substances could not be found in ETOX. 
Here we specify that only the \emph{'best'} match (in terms of the Levenshtein distance between query and results) is returned. 
A manual check confirms appropriate matches. 
Other options include: \emph{'all'} - returns all matches; \emph{'first'} - returns only the first match (not necessarily the best match); \emph{'ask'} - this enters an interactive mode, where the user is asked for a choice if multiple matches are found and \emph{'na'} which returns \code{NA} in case of multiple matches.

We use these data to retrieve basic information on the substances

<<etox_basic, message=FALSE>>=
etox_data <- etox_basic(ids$etoxid)
@

\pkg{webchem} always returns a named list (one entry for each substance) and the available output / information content can be very voluminous.
Therefore, we provide extractor functions for the common identifiers: CAS, SMILES and InChIKeys.
<<etox_cas, message=FALSE>>=
etox_cas <- cas(etox_data)
head(etox_cas)
@

A variety of data are available and we cannot provide extractor functions for each of those.
Therefore, if users need to extract other data, they have to write simple extractor functions (see following examples).

In the same manner we can now query other identifiers from other source using this CAS numbers (Fig. \ref{fig:fig1}):
<<token, echo=FALSE>>=
# this is a webchem specific token. 
# Please use it only for reproduction / testing.
token <- 'b44136c5-49cf-4df4-ba18-e4676449d1f4'
@

<<others, message = FALSE, warning=FALSE>>=
# query SMILES from PubChem
cids <- get_cid(etox_cas)
pc_data <- pc_prop(cids, properties = c('CanonicalSMILES'))
pc_smiles <- smiles(pc_data)

# query InChiKey from ChemSpider
# token <- '<yourtokenhere>'
csids <- get_csid(etox_cas, token = token)
cs_data <- cs_compinfo(csids, token = token)
cs_inchikey <- inchikey(cs_data)
@

<<table>>=
# combine in one data.frame
res <- data.frame(name = subs, cas = etox_cas, smiles = pc_smiles, 
                  cid = pc_data$CID, inchikey = cs_inchikey, 
                  csid = cs_data$csid, 
                  stringsAsFactors = FALSE)
@

Note, that in order to use the ChemSpider functions, a personal authentication key (\code{token}) is needed, which can be retrieved from the ChemSpider webpage. 
We end up with a compound table containing many different identifiers (Table~\ref{tab:comptable}), allowing easy identification and merging with other datasets, e.g. the lc50 dataset based on CAS.

\begin{table}[ht]
\centering
<<print_table, echo=FALSE, results = 'asis'>>=
tmp <- res[1:6, ]
tmp$smiles <- ifelse(is.na(tmp$smiles), NA, paste0(substr(tmp$smiles, 1, 7), '...'))
tmp$inchikey <- ifelse(is.na(tmp$inchikey), NA, paste0(substr(tmp$inchikey, 1, 7), '...'))
tmp[is.na(tmp)] <- '-'
colnames(tmp) <- c("Name", "CAS", "SMILES\\textsuperscript{a}","CID", "InChIKey\\textsuperscript{a}", "CSID")
print(xtable(tmp),
  floating = FALSE,
  hline.after = NULL,
  include.rownames = FALSE,
  add.to.row = list(pos = list(-1,0, 6),
                    command = c('\\toprule\n','\\midrule\n','\\bottomrule\n')),
  sanitize.text.function = function(x){x}
  )
@
\caption{Identifiers for the jagst datasets as queried with \pkg{webchem}. Only the first 6 entries are shown. - = not found. \textsuperscript{a} Only the first 7 characters are shown.}
\label{tab:comptable}
\end{table}

\subsection[Toxicity of different pesticide groups]{Toxicity of different pesticide groups}
Another question we might ask is \emph{How does toxicity vary between insecticide groups?}
Answering this question would require tedious lookup of insecticide groups for each of the 124 CAS numbers in the lc50 dataset.
The Compendium of Pesticide Common Names \citep{wood} contains such information and can be easily queried with \pkg{webchem}.

<<aw_query, message = FALSE>>=
# query the compendium using CAS-numbers
aw_data <- aw_query(lc50$cas, type = 'cas')
@

To extract the chemical group from the retrieved dataset, we need to write a simple extractor function and apply this to the retrieved data

<<subactivty, message = FALSE>>=
# shows internal structure of the data-object
# str(aw_data[[1]])
# extract chemical group
igroup <- sapply(aw_data, function(y) y$subactivity[1])
igroup[1:3]
@

Figure \ref{fig:fig2} displays the result after some more data-cleaning (see supplement for full code).
Overall, it took only 5 \proglang{R} statements to retrieve, clean and plot the data using \pkg{ggplot2} \citep{ggplot2}.

\begin{figure}[ht]
<<plot_lc50, echo=FALSE, out.width='0.7\\textwidth', fig.height=6>>=
# cleanup
lc50$type <- ifelse(grepl('carbamat', igroup), 'Carbamates',
              ifelse(grepl('neonicotinoid', igroup), 'Neo-\nnicotinoids', 
                ifelse(grepl('pyrethroid', igroup), 'Pyrethroids', 
                  ifelse(grepl('organophosphate', igroup), 'Organo-\nphosphates', 
                         'other'))))
lc50$type <- factor(lc50$type, levels = c("Pyrethroids", "Carbamates", "Organo-\nphosphates", "Neo-\nnicotinoids", "other"))

# plot
p <- ggplot(lc50, aes(x = type, y = value)) +
  geom_boxplot(fill = 'grey75') + 
  geom_jitter(width = 0.5) +
  scale_y_log10() +
  labs(y = expression(LC[50~'D.magna, 48h']), x = '') +
  theme_bw() +
  theme(text = element_text(size = 16))
p
@
\caption{Toxicity of different pesticide groups. LC50 values have been retrieved from EPA ECOTOX database, chemical groups from the Compendium of Pesticide Common Names \citep{wood}}
\label{fig:fig2}
\end{figure}

\subsection[Querying partitioning coefficients]{Querying partitioning coefficients}
Some data sources contain also data on chemical properties that can be queried.
Here we query for the lc50 data the $log~P_{oct/wat}$ from SRCs PHYSPROP database to build a simple Quantitative structure–activity relationship (QSAR) to predict toxicity.

<<qsar, message=FALSE, cache=TRUE, warning=FALSE>>=
# query PHYSPROP DATABASE
pp_data <- pp_query(lc50$cas)
@

The database contains predicted and experimental values.
Extracting $log~P_{oct/wat}$ from the data object is slightly more complicated,  
because i) for some compounds no data could be found and ii) the data-object has a more complex structure (a data frame within a list).

<<qsar2, warnings = FALSE>>=
# look  at internal structure
# str(pp_data[[1]])
lc50$logp <- sapply(pp_data, function(y){
  if (length(y) == 1 && is.na(y))
    return(NA)
  y$prop$value[y$prop$variable == 'Log P (octanol-water)']
})
@

We opted for this more complex approach, because the information available is very divers and we cannot provide an extractor function for each purpose.
Moreover, it gives users full control on how to organize their data. 
Nevertheless, we provide in the documentation of each function how to extract more complicated parts of the data.
The resulting data and model is displayed in Figure~\ref{fig:fig3}.

\begin{figure}[ht]
<<plot_qsar, echo=FALSE, out.width='0.7\\textwidth', fig.height=6, warning=FALSE>>=
# model
mod <- lm(log10(value) ~ logp, data = lc50)
coefs <- round(coef(mod), 2) # coeficients
rmse <- round(summary(mod)$sigma, 2) # RMSE

# check data types
logp_type <- sapply(pp_data, function(y){
  if (length(y) == 1 && is.na(y))
    return(NA)
  y$prop$type[y$prop$variable == 'Log P (octanol-water)']
})
tabs <- table(logp_type)
sna <- sum(is.na(logp_type))

p <- ggplot(lc50, aes(x = logp, y = log10(value))) +
  geom_point() +
  geom_smooth(method = 'lm') +
  labs(x = expression(log~P[oct/wat]), y = expression(log~LC[50])) +
  theme_bw()
p
@
\caption{Simple QSAR for predicting logLC50 of pesticides by log P. 
Log P values have been retrieved from SRC Physprop database (\Sexpr{tabs[2]} experimental data, \Sexpr{tabs[1]} estimated data and \Sexpr{sna} substances without data). 
Blue line indicates the regression model ($log LC_{50} = \Sexpr{coefs[1]}\Sexpr{coefs[2]} logP$, RMSE = \Sexpr{rmse}).}
\label{fig:fig3}
\end{figure}

\subsection[Utility functions]{Utility functions}
\pkg{webchem} provides also basic functions to check identifiers that can be used for data quality assessment.
The functions either use simple formatting rules or web resources.

<<check>>=
# simple formatting check
is.inchikey('BQJCRHHNABKAKU-KBQPJGBKS-AN')
# or using ChemSpider
is.inchikey('BQJCRHHNABKAKU-KBQPJGBKSA-5', 
      type = 'chemspider')
# formatting check
is.cas('64-17-6')
@

\section[Discussion]{Discussion}
\subsection[Related software]{Related software}
Within the \proglang{R} ecosystem , there are only few similar projects:
\pkg{rpubchem} \citep{rpubchem_2014} and \pkg{ChemmineR} \citep{chemminer_2008} both provide an interface to PubChem that is more extensive than the current one in \pkg{webchem}.
\pkg{WikidataR} \citep{wikidatar_2016} provides an interface to wikidata that could be used to retrieve chemical data from Wikipedia.
However, it does not provide predefined methods for chemical data like \pkg{webchem}.
Within the \proglang{Python} ecosystem (a popular programming language) the libraries \pkg{PubChempy} \citep{pubchempy}, \pkg{ChemSpiPy} \citep{chemspipy} and \pkg{CIRpy} \citep{cirpy} are available for similar tasks as those outlined here.
\pkg{webchem} is not specialized and tries to integrate many data sources and for some of these provides a unique programmatic interface.
The Chemical Translation Service \citep{Wohlgemuth_Haldiya_Willighagen_Kind_Fiehn_2010}, which is also one of the sources that can be queried,  allows batch conversion of chemical identifiers.
However, it does not provide access to other data (experimental, modeled or legal data).

\subsection[Open Science]{Open Science}
More and more scientific data are becoming publicly available \citep{Gewin_2016, Reichman_Jones_Schildhauer_2011,Boyle_Guha_2011}, either in public data repositories or as supplement to publications.
To be usable for other researchers chemical compounds should be properly identified, not only by chemical names but also with accompanying identifiers like InChIKey, SMILES and authority-assigned identifiers.
\pkg{webchem} provides an easy way to create such meta tables like Table \ref{tab:comptable}.

\pkg{webchem} facilitates chemical data availability to researchers.
However, data quality is also crucial for data-analysis \citep{Stieger_2014}. 
Ensuring good quality requires additional effort and methods to be developed.

\subsection[Further development]{Further development}
We have outlined only a few use cases that likely will be useful for many researchers.
Given the huge amount of publicly available information, many other possibilities can be envisioned.
\pkg{webchem} is currently under active development and several other data sources have not been implemented yet but may be in the future.
GitHub makes contributing easy and we strongly encourage contribution to the package.
Moreover, comments, feedback and feature requests are highly welcome.

\section[Conclusions]{Conclusions}
Researchers need to have easy access to global knowledge on chemicals.
\pkg{webchem} can save \emph{"hundreds of working hours"} gathering this knowledge \citep{Munch_Galizia_2016}, so that researchers can focus on other tasks.

\section*{Acknowledgements}
We thank all resource maintainers for their work making their data open to the public.
We thank Johannes Ranke and Daniel Münch for their contributions to the webchem package, as well as all users who provided feedback and feature requests.

\bibliography{references.bib}

\end{document}